---
title: "Summary analysis for Jan 2021 meeting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
library(ggridges)
library(patchwork)
library(growthrates)
library(ggpubr)
#library(rayshader)
library(plotly)
library(knitr)
library(PerformanceAnalytics)
library(car)
library(MASS)
library(vegan)
library(lubridate)
library(scales)
library(lme4)
library(nlme)

theme_set(theme_bw()+
            theme(axis.text=element_text(size=12),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16)))
```



# Setup 

Read in data and check out how many observations of each variable we have to play with

```{r}
sisyn <- read_csv("20201111_masterdata_RAW.csv")

sisyn %>% count(variable, sort = TRUE) %>% kable()
```

Spread the data into wide format, make year column, and summarize annual averages into a data table

```{r}
# Spreading data to give each variable a column, averaging values by date/site

sisyn_wide <- sisyn %>% 
  mutate(sample=paste(LTER, site, Sampling.Date, sep="_")) %>%
  pivot_wider(names_from = variable, values_from = value, values_fn = mean, id_cols = c(sample, LTER, Sampling.Date, site, `Site/Stream.Name`))

# making extracting year into its own column

sisyn_wide <- sisyn_wide %>% 
  mutate(sample.date=date(Sampling.Date)) %>%  
  mutate(year=year(sample.date))

# making another dataset of just annual averages for DSi, NOx, and PO4

sisyn_annual <- sisyn_wide %>% 
  group_by(year, LTER, site) %>% 
  summarize(DSi = mean(DSi, na.rm=TRUE),
            NOx = mean(NOx, na.rm=TRUE),
            PO4 = mean(PO4, na.rm=TRUE)) %>% 
  mutate(site = str_replace_all(site, " ", "_"))

# some have NAs for DSi so those should be removed
i <- which(is.na(sisyn_annual$DSi))
sisyn_annual[i,]

sisyn_annual <- sisyn_annual %>% 
  filter(!is.na(DSi))

i <- which(is.na(sisyn_annual$DSi))
sisyn_annual[i,]

#making each column contain DSi conc for each site, since this is the only way I could figure out how to do >400 regressions at once

sisyn_annual_wide <- sisyn_annual[1:4]
sisyn_annual_wide <- sisyn_annual_wide %>% 
  pivot_wider(names_from = site, values_from = DSi)


sisyn_annual %>% 
  filter(!is.na(DSi)) %>%
  count(site)

sisyn_annual <-  sisyn_annual %>% 
  filter(!is.na(DSi)) %>%
  group_by(LTER, site) %>% 
  count(site, sort = TRUE) %>% 
  kable()
```

Some sites only have one observation so we need to remove those for regressions to work

```{r}
sisyn_annual %>% 
  filter(!is.na(DSi)) %>%
  group_by(LTER, site) %>% 
  count(site, sort = TRUE) %>% 
  kable()



```



# DSi concentrations by year and site


```{r}

ggplot(sisyn_annual, aes(year, DSi, group=site))+
  geom_smooth(method="lm", se=FALSE)+
  scale_y_log10()+
  facet_wrap(~LTER, scales = "free")

ggplot(sisyn_annual, aes(year, DSi, group=site))+
  geom_point()+
  geom_line()+
  scale_y_log10()+
  facet_wrap(~LTER, scales = "free")

```


```{r}

n <- 435

my_lms <- lapply(3:n, function(x) lm(as.matrix(sisyn_annual_wide[,x]) ~ year, data=sisyn_annual_wide))
                   
coefs <- sapply(my_lms, coef)
coefs <- data.frame(coefs)
#rownames(coef)[1:435] <- myvars #FIGURE OUT HOW TO DO THIS!!!!

summaries <- lapply(my_lms, summary)
rquares <- sapply(summaries, function(x) c(r_sq = x$r.squared, 
                                adj_r_sq = x$adj.r.squared))

pvals <- lapply(summaries, function(x) x$coefficients[, c(1,4)])

coefs <- rownames_to_column(coefs) %>% tibble

rownames(coefs)[1] <- "coef_type"

coefs_long <- coefs %>% 
  pivot_longer(!rowname,
               names_to = c("coeftype"),
               values_to="value")
               
               
names_to = c("incercept", "slope"),
   names_sep = 


anscombe   
anscombe %>%
 pivot_longer(everything(),
   names_to = c(".value", "set"),
   names_pattern = "(.)(.)")

```



trying LMER approach to get slopes for each site

```{r}

sisyn_annual <- sisyn_annual %>% 
  mutate(site=paste(LTER, site, sep="_"))

fit1 <- lmer(DSi ~ year + (1+year|site), data=sisyn_annual)
summary(fit1)
anova(fit1)
ranef(fit1)


plot(fit1)

```

trying lapply and shit

```{r}




GSWS09 <- filter(sisyn_annual, site=='GSWS09')
fit <- lm(DSi~year, data=GSWS09)
summary(fit)
plot(fit)


set.seed(1)
# run n regressions

n <- 438

my_lms <- sisyn_annual_wide %>%
  lapply(1:n, function(x) lm(DSi[,x] ~ year[,x]))

# extract just coefficients
sapply(my_lms, coef)

```


```{r}

mod_summaries <- list()                  # Create empty list


for(i in 3:ncol(sisyn_annual_wide)) {                 # Head of for-loop
 
  sites_i <- colnames(sisyn_annual_wide)[3:i]    # Create vector of predictor names
  mod_summaries[[i - 1]] <- summary(     # Store regression model summary in list
    lm(year ~ ., data=sisyn_annual_wide[ , c("year", sites_i)]))
 
}

set.seed(98274)                          # Creating example data
y <- rnorm(1000)
x1 <- rnorm(1000) + 0.2 * y
x2 <- rnorm(1000) + 0.2 * x1 + 0.1 * y
x3 <- rnorm(1000) - 0.1 * x1 + 0.3 * x2 - 0.3 * y
data <- data.frame(y, x1, x2, x3)
head(data)  

mod_summaries <- list()                  # Create empty list

for(i in 2:ncol(data)) {                 # Head of for-loop
 
  predictors_i <- colnames(data)[2:i]    # Create vector of predictor names
  mod_summaries[[i - 1]] <- summary(     # Store regression model summary in list
    lm(y ~ ., data[ , c("y", predictors_i)]))$coefficients
 
}

mod_summaries
```


```{r}

set.seed(1)

# number of columns in the Lung and Blood data.frames. 22,000 for you?
n <- 20

myvars <- as.list(colnames(sisyn_annual_wide[3:436]))


# run n regressions
my_lms <- lapply(myvars, function(x) lm(sisyn_annual_wide[,x] ~ year,
                                     data=sisyn_annual_wide))

my_lms <- lapply(myvars, function(x) lm(as.matrix(sisyn_annual_wide[,x]) ~ year, data=sisyn_annual_wide))
                   
sapply(my_lms, coef)

lms <- function(x) {
  fit <- lm(sisyn_annual_wide[,x] ~ year, data=sisyn_annual_wide)
  return(summary(fit))
}

lms(3)

fit <- lm(GSWS09~year, data=sisyn_annual_wide)
summary(fit)

fit <- lm(as.matrix(sisyn_annual_wide[,3])~year, data=sisyn_annual_wide)
summary(fit)
                                     
# extract just coefficients
sapply(my_lms, coef)

# if you need more info, get full summary call. now you can get whatever, like:
summaries <- lapply(my_lms, summary)
# ...coefficents with p values:
lapply(summaries, function(x) x$coefficients[, c(1,4)])
# ...or r-squared values
sapply(summaries, function(x) c(r_sq = x$r.squared, 
                                adj_r_sq = x$adj.r.squared))
```


```{r}



myvars <- as.list(colnames(sisyn_annual_wide[3:436]))


fit <- lapply(myvars, function(year, myvars)
    lm(year ~ myvars, data=sisyn_annual_wide))

```


```{r}
```


```{r}
```

